---
title: "Mith52_Sujit"
author: "Sujit"
date: "5 January 2019"
output: html_document
---

```{r}
rm(list=ls(all=T))
```

##Load libraries

```{r}
library(tidyr)
library(data.table)
library(lubridate)
library(eeptools)
library(dplyr)
```

##Read the train files and analyse it

```{r}
demographics = read.csv("Train_Demographics.csv",header=TRUE,sep=",")
transactions = read.csv("Train_Transactions.csv",header=TRUE,sep=",")
customer = read.csv("Train.csv",header=TRUE,sep=",")
head(transactions)
```

```{r}
demographics = demographics[order(demographics$CustomerId),]
transactions = transactions[order(transactions$CustomerId),]
customer = customer[order(customer$CustomerId),]
```

```{r}
nrow(transactions)
sum(is.na(transactions))
sum(is.na(demographics))
sum(is.na(customer))
```
```{r}
str(transactions)
str(demographics)
str(customer)
#transactions$CustomerId = as.character(transactions$CustomerId)
#demographics$CustomerId = as.character(demographics$CustomerId)
#customer$CustomerId = as.character(customer$CustomerId)
```

##Feature engineering

```{r}
transactions$amount <- transactions$Quantity*transactions$UnitPrice

head(transactions)
```

```{r}
tran_amount = transactions %>% 
  group_by(CustomerId) %>% 
  summarise(noVisits = n(),total_am = sum(amount)) 

tail(tran_amount)
nrow(tran_amount)
```

```{r}
tran_a = transactions %>% 
  group_by(CustomerId,BillNumber) %>% 
  summarise(total_am = sum(amount))

tran_a %>% 
  group_by(CustomerId) %>% 
  summarise(noVisits = n(),total_amt = sum(total_am))
```


## merge the data sets

```{r}
merge_demo_tran = merge(demographics,tran_amount,by.x="CustomerId",by.y = "CustomerId")
head(merge_demo_tran)
merged_customer = merge(merge_demo_tran,customer,by.x="CustomerId",by.y = "CustomerId")
head(merged_customer)
nrow(merged_customer)
```

```{r}
str(merged_customer)
```


```{r}
sum(is.na(merged_customer))
```


```{r}
merged_customer$Familysize = as.numeric(merged_customer$Familysize)
merged_customer$noVisits = as.numeric(merged_customer$noVisits)
merged_customer$DOB = as.Date(merged_customer$DOB)
merged_customer$age=floor(age_calc(merged_customer$DOB, units = "years"))

str(merged_customer)
head(merged_customer)
```

## drop customer id and DOB columns
```{r}
drop_cols <- c("CustomerId","DOB")
merged_customer[ ,drop_cols] <- list(NULL)

str(merged_customer)
```


```{r}
cat_cols = c("gender","MaritalStatus","CustomerValue")
num_cols = setdiff(colnames(merged_customer),cat_cols)
cat_cols
num_cols
```

## Create data analysis report

```{r}
#install.packages("DataExplorer")
library(DataExplorer)
create_report(merged_customer)
```

```{r}
plot(merged_customer$noVisits,merged_customer$CustomerValue,xlab="No of Visits",ylab="Value of customer",main="CustomerValue vs NoOfVisits", col= "blue")

plot(merged_customer$total_am,merged_customer$CustomerValue,xlab="Total Spend",ylab="Value of customer",main="TotalSpend vs NoOfVisits", col= "blue")
```

```{r}
library(ggplot2)
ggplot(data = merged_customer, aes(x = CustomerValue)) + geom_bar(fill = 'blue')
```


```{r}
ggplot(data = merged_customer, aes(x = age, fill = CustomerValue)) + geom_histogram()
```

```{r}
ggplot(data = merged_customer, aes(x = total_am, fill = CustomerValue)) + geom_histogram()
```

```{r}
ggplot(data = merged_customer, aes(x = noVisits, fill = CustomerValue)) + geom_histogram()
```


```{r}
ggplot(data = merged_customer, aes(x = Familysize, fill = CustomerValue)) + geom_histogram()
```


```{r}
ggplot(data = merged_customer, aes(x = gender, fill = CustomerValue)) +
  geom_bar()
```

```{r}
ggplot(data = merged_customer, aes(x = MaritalStatus, fill = CustomerValue)) +
  geom_bar() 
```



## draw coorelation plot

```{r fig.height= 4, fig.width = 5}
library(corrplot)

corrplot(cor(merged_customer[,num_cols], use = "complete.obs"))
```

## split train and test on the merged dataset

```{r}

library(caret)

set.seed(007)

trainIndex <- createDataPartition(merged_customer$CustomerValue, p = .7, list = F)

train_data <- merged_customer[trainIndex, ]

validate_data <- merged_customer[-trainIndex, ]
```


##build Decision tree and analyse the performance
```{r}
library(C50)
c5_tree <- C5.0(CustomerValue ~ . , train_data)
```

```{r}
C5imp(c5_tree, metric = "usage")
```

```{r}
summary(c5_tree)
```

```{r, fig.width= 35, fig.height=15}

plot(c5_tree)
```

```{r}
preds <- predict(c5_tree, validate_data)
```

```{r}
library(caret)

confusionMatrix(preds, validate_data$CustomerValue)
```

```{r}
preds=as.vector(preds)
table(preds)
```


## Build multinomial regression and verify accuracy
```{r}
library(nnet)
multi_class = multinom(CustomerValue ~ ., train_data)
multi_preds = predict(multi_class , validate_data ,type = "class")
confusionMatrix(multi_preds, validate_data$CustomerValue)
```

```{r}
multi_preds=as.vector(multi_preds)
table(multi_preds)
```

##Build a Random Forest model and verify accuracy
```{r}
library(randomForest)
RFModel = randomForest(CustomerValue ~ ., data = train_data, keep.forest = TRUE, ntree=200)
print(RFModel)
```

```{r}
RFModel$importance
round(importance(RFModel),2)
varImpPlot(RFModel)
```

```{r}
preds_rf <- predict(RFModel, validate_data)
confusionMatrix(preds_rf, validate_data$CustomerValue)
```

## dummify categorical variables for XGBoost model

```{r}
dummies <- dummyVars(~ gender +  MaritalStatus, merged_customer)
df_all_ohe <- as.data.frame(predict(dummies, merged_customer))
merged_customer <- cbind(merged_customer[,!(names(merged_customer) %in%
                                           c('gender','MaritalStatus'))],df_all_ohe)
str(merged_customer)

```


```{r}
library(caret)

set.seed(009)

xg_trainIndex <- createDataPartition(merged_customer$CustomerValue, p = .7, list = F)

xg_train_data <- merged_customer[xg_trainIndex, ]

xg_validate_data <- merged_customer[-xg_trainIndex, ]

CustomerValue <- recode(xg_train_data$CustomerValue,'High'=0, 'Medium'=1, 'Low'=2)

validate_CustomerValue <- recode(xg_validate_data$CustomerValue,'High'=0, 'Medium'=1, 'Low'=2)

xg_train_data$CustomerValue<-NULL
xg_validate_data$CustomerValue<-NULL
str(xg_validate_data)
```

## build XGBoost model
```{r}
library(xgboost)
xgb <- xgboost(data = data.matrix(xg_train_data), 
 label = CustomerValue, 
 eta = 0.1,
 max_depth = 15, 
 nround=25, 
 subsample = 0.5,
 colsample_bytree = 0.5,
 seed = 1,
 eval_metric = "merror",
 objective = "multi:softmax",
 num_class = 3,
 nthread = 3
)
```

```{r}
xgb$feature_names
```

## verify the accuracy
```{r}
xgb_pred <- predict(xgb, data.matrix(xg_validate_data))
#xgb_pred <- recode(as.character(xgb_pred),'0'='High', '1'='Medium', '2'='Low')

#xg_validate_data <- cbind(xg_validate_data,validate_CustomerValue)
confusionMatrix(as.factor(xgb_pred), as.factor(validate_CustomerValue))
```


```{r}
mat <- xgb.importance (feature_names = colnames(xg_train_data),model = xgb)
xgb.plot.importance (importance_matrix = mat[1:8]) 
```


## run the same data analysis steps for test data
```{r}
test_demographics = read.csv("Test_Demographics.csv",header=TRUE,sep=",")
test_transactions = read.csv("Test_Transactions.csv",header=TRUE,sep=",")
test_customer = read.csv("Test.csv",header=TRUE,sep=",")
test_demographics = test_demographics[order(test_demographics$CustomerId),]
test_transactions = test_transactions[order(test_transactions$CustomerId),]
test_customer = data.frame(test_customer[order(test_customer$CustomerId),])
sum(is.na(test_demographics))
sum(is.na(test_transactions))
sum(is.na(test_customer))
```

```{r}
test_transactions$amount <- test_transactions$Quantity*test_transactions$UnitPrice
test_tran_amount = test_transactions %>% 
  group_by(CustomerId) %>% 
  summarise(noVisits = n(),total_am = sum(amount)) 
test_merged_customer = merge(test_tran_amount,test_demographics,by.x="CustomerId",by.y =
                               "CustomerId")
test_merged_customer$Familysize = as.numeric(test_merged_customer$Familysize)
test_merged_customer$noVisits = as.numeric(test_merged_customer$noVisits)
test_merged_customer$DOB = as.Date(test_merged_customer$DOB)
test_merged_customer$age=floor(age_calc(test_merged_customer$DOB, units = "years"))
drop_cols <- c("CustomerId","DOB")
test_merged_customer[ ,drop_cols] <- list(NULL)

str(test_merged_customer)
```

```{r}
test_df_all_ohe <- as.data.frame(predict(dummies, test_merged_customer))
test_merged_customer <- cbind(test_merged_customer[,!(names(test_merged_customer) %in%
                                           c('gender','MaritalStatus'))],test_df_all_ohe)
str(test_merged_customer)
```
```{r}
test_merged_customer <- test_merged_customer[c("Familysize","noVisits","total_am","age",
                                             "gender.F","gender.M","MaritalStatus.No",
                                             "MaritalStatus.Yes")]

str(test_merged_customer)
```


## Get the test predictions for XGBoost model
```{r}
test_xgb_pred <- predict(xgb, data.matrix(test_merged_customer))
test_xgb_pred
test_xgb_pred <- recode(as.character(test_xgb_pred),'0'='High', '1'='Medium', '2'='Low')
test_xgb_pred

final_val = cbind(test_demographics , test_xgb_pred)
final_val
xgb_submission = data.frame('CustomerId'=final_val$CustomerId , 
                           'CustomerValue' = final_val$test_xgb_pred)
write.csv(xgb_submission , file = "submission.csv" ,row.names = FALSE)
```


## Get the test predictions for Random Forest model
```{r}
#test_preds_rf <- predict(RFModel, test_merged_customer)

#final_val = cbind(test_demographics , test_preds_rf)
#final_val
#rf_submission = data.frame('CustomerId'=final_val$CustomerId , 
#                           'CustomerValue' = final_val$test_preds_rf)
#write.csv(rf_submission , file = "submission.csv" ,row.names = FALSE)
```


















